---
title: "Data Visualization - Case Study 5 to 7"
author: "Srushti Padade"
date: "April 11, 2020"
output:
  word_document: default
  html_document: default
---

```{r}
library(ggplot2)
```

#CLASS PRACTICE 5

1.	Plot the histogram using ggplot()

```{r}
ggplot(mpg, aes(hwy)) +
  geom_histogram()
```

2.	Plot the distribution of the variables using geom_freqpoly()

```{r}
ggplot(mpg, aes(hwy)) +
  geom_freqpoly()
```

3.	Plot box plot using geom_boxplot()

```{r}
ggplot(mpg)+ 
  geom_boxplot(aes(x = displ, y = cty))
```

4.	What is the purpose of Histograms and Density plot?

Answer:
A Density Plot visualises the distribution of data over a continuous interval or time period. This chart is a variation of a Histogram that uses kernel smoothing to plot values, allowing for smoother distributions by smoothing out the noise. This requires using a density scale for the vertical axis.

A histogram can be used to compare the data distribution to a theoretical model, such as a normal distribution.

5.	Name another Univariate plot?

Answer:
Univearite plots are :
- Histogram
- Density Plot
- Bar plot

```{r}
ggplot(mpg, aes(displ)) + 
  geom_bar()
```

6.	Plot the following using ggplot() + facet_wrap()

```{r}
ggplot(mpg, aes(cty, fill = drv)) +
  geom_histogram(binwidth = 0.5) +
  facet_wrap(~drv, ncol = 1)
```

7.	Plot the following graph:

```{r}
ggplot(mpg, aes(displ,color = class)) + geom_freqpoly(binwidth = 0.5)
```

CLASS PRACTICE 6 - TIME SERIES

1.	Use the "economics" dataset and Plot the following using geom_line():

```{r}
ggplot(economics, aes(date, pop)) + 
  geom_line()
```

2.	Plot the unemployment growth rate over a period of time:

```{r}
ggplot(economics, aes(date, unemploy)) +  geom_line()
```

3.	Plot the graph to show how long people were unemployed?

```{r}
ggplot(economics, aes(date, unemploy / pop)) +
  geom_line()
```

OR
```{r}
ggplot(economics, aes(date, uempmed)) + geom_line()
```

4.	Plot the below graph(HINT: Use POSIXlt() ; ggplot() +  geom_path() + geom_point()

```{r}
year <- function(x) as.POSIXlt(x)$year + 1900

ggplot(economics, aes(unemploy/pop, uempmed)) +
  geom_path() + 
  geom_point(aes(color = year(date)))
```

#CLASS PRACTICE 7

1.	Explore the distribution of the "carat" variable in the diamonds dataset. What banwidth reveals the most interesting patterns?

```{r}
ggplot(diamonds) +
  geom_histogram(aes(carat), binwidth = 0.05)
```

```{r}
ggplot(diamonds) +
  geom_histogram(aes(carat), binwidth = 0.1)
```

```{r}
ggplot(diamonds) +
  geom_histogram(aes(carat), binwidth = 0.3)
```

The binwidth function for the histogram help us have a closer look at the distribution of the caret for diamonds.

Bin selection will play a significant role in visualizations, with a possible change in frequency readouts and shape of the curve or function.

We can see the data is left sweked and have negative relation. Smaller carat diamonds have higher sale and are easier to obtain.

Bigger rocks are hard to come by.

```{r}
ggplot(diamonds, aes(x=carat)) + 
  geom_freqpoly() + 
  ggtitle("Diamond Frequency by Carat") + 
  xlab("Carat Size") + 
  ylab("Count")
```

2.	Explore the distribution of the "price" variable in the diamonds data. How does the distribution vary by cut?

```{r}
ggplot(diamonds) +
  geom_histogram(aes(x = price), binwidth = 500) +
  facet_grid(cut~.)
```

```{r}
ggplot(diamonds) +
  geom_boxplot(aes(x = cut, y = price))
```

As we can see, the diamonds cut of all types carry prices of all types, not really a way to determine how good or expensive a diamond is.

Thus the diamonds price cannot be determined based on only one characteristics.

```{r}
ggplot(diamonds) + 
  geom_histogram(binwidth=100, aes(price)) +
  ggtitle("Diamond Price Distribution by Cut") + 
  theme_minimal() + 
  facet_wrap(~cut, scales="free_y")
```

Above plot thus provides us an average picture of price distribution of diamonds based on there cuts.