---
title: "Data Visualization - Case Study 1 to 4"
author: "Srushti Padade"
date: "April 7, 2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

#CLASS PRACTICE 1

```{r}
library(ggplot2)
library(tidyverse)
options(warn=-1)
```

1. Plot the below Graph:

```{r}
ggplot(mpg, aes(x=displ, y = hwy)) + 
  geom_point()
```

2. How would you describe the relationship between cty and hwy (plot graph using ggplot)? Do you have any concerns about drawing conclusions from that plot?

```{r}
ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
```

Answer:
The 'cty' and 'hwy' variables describe the fuel economy. By looking in the graph we can see a clear positive Linear relationship between them.
Generally, cars are more fuel efficient on highways rather than in cities due to inbuild speed functionality.
However here We can infer that, cars that are fuel efficient both on highway as well as city roads.
Drawing this conclusion was easy using the above plot as the clear picture could be seen.

3. Plot the following graph and explain the output:

```{r}
ggplot(mpg, aes(model, manufacturer)) + 
  geom_point()
```

Answer:
The plot represents relation between each mactufacturer with the car models. By looking at the scattered plot no defined conclusion could be drawn.

4. Is the above graph useful? How could you modify the data to make it more informative?

Answer:
Since the original plot could not be used to infer and particular relation ship between the manufacturer and there models.

Thus we have tried to  manipulate the manufacturer to its models relation to understand the rate at which the Fuel consumption for each pair of Manufacturer and the model.

```{r}
Combine_ManufacturerToModel <- mpg %>% 
  transmute("ManufacturerToModel" = paste(manufacturer, model, sep = " - "))

ggplot(Combine_ManufacturerToModel, aes(ManufacturerToModel)) +
  geom_bar() + coord_flip()
```

#CLASS PRACTICE 2

1. Plot the following using ggplot()
```{r}
ggplot(mpg, aes(x=displ, y = cty, colour = class)) +
  geom_point()
```

2.	Plot the following (adding one more dimension):

```{r}
ggplot(mpg, aes(x=displ, y = hwy, shape = drv)) + 
  	geom_point(colour = "orange")
```

3.	Plot the following graph (adding more dimensions)

```{r}
ggplot(mpg, aes(x=displ, y = cty, size = drv)) + 
 geom_point(colour="green")
```

4.	Plot the following graph:

```{r}
ggplot(mpg, aes(x=displ, y = hwy, shape = drv, color = class)) + 
  geom_point()
```

5.	Plot the following graph:

```{r}
ggplot(mpg, aes(x=displ, y = hwy, shape = drv, color = class, size = cyl)) + 
  geom_point()
```

6. How is drive train related to fuel economy? Plot and show

```{r}
#Fuel economy in CITY
ggplot(mpg, aes(drv, cty)) + 
  geom_boxplot() +
  scale_x_discrete(limits = c("f", "r", "4"),
                   labels = c("Front wheel", "Rear wheel", "Four wheel"))
#fuel economy in HIGHWAY
ggplot(mpg, aes(drv, hwy)) + 
  geom_boxplot() +
  scale_x_discrete(limits = c("f", "r", "4"),
                   labels = c("Front wheel", "Rear wheel", "Four wheel"))
```

Answer:

Front wheeler vehicles are more fuel efficient for both highway and City drives.

7.	How is drive train related to engine size and class? Plot and show

```{r}
ggplot(mpg, aes(class, displ, color = drv)) + geom_point()
```

Answer:
- From the above plot we can say that cars of class 'suv', pickup have large engine size and most of them are four wheel drive.
- Cars of Class 'compact', 'subcompact', 'midsize', and 'minivan' have comparetively less engine size and most of them are front wheel drive.
- 'Two seater' cars have large engine size than any other and all of them are rear wheel drive.

8.	What happens if you map trans variable to shape? Why?

```{r}
ggplot(mpg, aes(cty, hwy, shape = trans)) + 
  geom_point()
```

Answer:
The shape palette can deal with a maximum of 6 discrete values because more than 6 becomes difficult to discriminate; you have 10. Consider specifying shapes manually if you must have them.
Removed 96 rows containing missing values (geom_point).
Hence loss of our data.

#CLASS PRACTICE 3

1. Plot this graph using ggplot()

```{r}
ggplot(mpg, aes(displ,cty)) +
geom_point(colour="orange")  + facet_wrap(~class)
```

2. Use facetting to explore the 3-way relationship between fuel economy, engine size, and number of cylinders. 

```{r}
#For cities
ggplot(mpg, aes(displ, cty)) + 
  geom_point() +
  facet_wrap(~ cyl)
```

```{r}
#For highways
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() +
  facet_wrap(~ cyl)
```

3.	How does facetting by number of cylinders change your assessment of the relationship between engine size and fuel economy?

Answer:
- 4 cylinders the relationship between fuel economy and engine size is negatively related.
- 5 cylinders we cannot intrepret any relation 
- 6 cylinders there is no much change in fuel economy with the increase of engine size but still we can say that is negatively related.
- 8 cylinders it is same as 6 cylinders but this is positively related.

#CLASS PRACTICE 4

1. Plot the following graph (adding smooth()):

```{r}
ggplot(mpg, aes(x=displ, y = cty)) +
  geom_point(colour="green") + 
  geom_smooth()
```

2. Change the method to linear (method="lm")

```{r}
ggplot(mpg, aes(x=displ, y = cty)) + 
  geom_point(colour="green") +
  geom_smooth(method = "lm")
```

3. Changing the span parameter to control wiggliness of the line:

```{r}
ggplot(mpg, aes(x=displ, y = cty)) +
  geom_point(colour="red") + 
  geom_smooth(span=0.4)
```

4. How is this smooth() function useful in interpreting results?

Answer:
It can be hard to view trends with just points alone. Many times we wish to add a smoothing line in order to see what the trends look like. This can be especially helpful when trying to understand regressions.
It aids the eye in seeing patterns in the presence of overplotting.

5.	Can you use this to explain this to a senior management executive

Answer: 
The normal plot shoews the scatter plot of the data and is difficut to interpret. As a senior management the smoothened line provided by the smooth function, it is easier to get a clear picture to the data and what inference that could be taken out of it.

6.	Try changing span parameter, Values range from 0 to 1

```{r}
ggplot(mpg, aes(x=displ, y = cty)) +
  geom_point(colour="red") + 
  geom_smooth(span=0.1)
```

```{r}
ggplot(mpg, aes(x=displ, y = cty)) +
  geom_point(colour="red") + 
  geom_smooth(span=1)
```

Answer:
When we changed the span = 1 we can see that this is much smoother. The span can be varied from 0 to 1, where 0 is very rough and 1 is very smooth.