---
title: "Data Visualization - Case Study 8 to 9"
author: "Srushti Padade"
date: "April 13, 2020"
output:
  html_document: default
  word_document: default
---

```{r}
library(ggplot2)
library(dplyr)
```

#CLASS PRACTICE - 8

1.	 Plot the following graph and label as shown:

```{r}
ggplot(economics, aes(date, unemploy)) +
  geom_line() + 
  labs(title = "Unemployment Rates",  y="Unemployment",  x="Year", caption = "Unemployment Rates")
```

2.	Execute the following program to annotate the graph:

```{r}
p <- ggplot(mpg, aes(displ, cty)) + 
  geom_point(data = filter(mpg, manufacturer == "audi"), colour = "red", size = 3) + 
  geom_point() 

p + 
  annotate(geom = "curve", x = 3, y = 35, xend = 2.8, yend = 20, curvature = .3, arrow = arrow(length = unit(2, "mm"))) +
  annotate(geom = "text", x = 3.1, y = 35, label = "Audi", hjust = "left")
```

```{r}
#install.packages("ggforce")
library(ggforce)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() + 
  geom_mark_ellipse(aes(label = cyl, group = cyl))
```

```{r}
#install.packages("gghighlight")
library("gghighlight")

ggplot(mpg, aes(displ, cty, colour = factor(drv))) +
  geom_point() + 
  gghighlight::gghighlight() + 
  facet_wrap(vars(drv))
```

3.	Try following plots:

```{r}
#Plot 1:
  ggplot(mpg, aes( displ,cty)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
#Plot 2:
  ggplot(mpg, aes(cty,displ)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
#Plot 3:
  ggplot(mpg, aes(displ, cty)) + 
  geom_point() + 
  geom_smooth() + 
  coord_flip()
```

4.	Using "Annotation" discussed earlier, highlight car type "Honda"

```{r}
q <- ggplot(mpg, aes(displ, cty)) + 
  geom_point(data = filter(mpg, manufacturer == "Honda"), colour = "orange", size = 3) + 
  geom_point() 

  
q + 
  annotate(geom = "curve", x = 3, y = 30, xend = 2, yend = 27, curvature = .3, arrow = arrow(length = unit(2, "mm"))) +
  annotate(geom = "text", x = 3.1, y = 30, label = "Honda", hjust = "left")
```

5.	Using ggforce() to highlight the "Class" of cars on "City" mileage based on engine (displ)

```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point() + 
  geom_mark_ellipse(aes(label = class, group = class))
```
  
6.	Using gghighlight() cluster "transmission" type

```{r}
ggplot(mpg, aes(displ, cty, colour = factor(trans))) +
  geom_point() + 
  gghighlight() + 
  facet_wrap(vars(trans))
```

#CLASS PRACTICE 9

1. Execute the following commands to build the first layer:

```{r}
a <-ggplot(mpg, aes(displ, hwy))
a
```

2.  Add aesthetics to the plot:

```{r}
b <- a + 
  layer(
    mapping = NULL, 
    data = NULL,
    geom = "point", 
    stat = "identity",
    position = "identity"
  )

b
```

3.  Can you change "colour" of points?

```{r}
b +
  geom_point(color = "red")
```

4.  Fit the line to the data points by executing following ggplot() program:

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  theme(legend.position = "none")
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = class)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(legend.position = "none")
``` 

4. Add two different type of line "methods" by executing following program:

```{r} 
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() +
  geom_smooth(aes(colour = "loess"), method = "loess", se = FALSE) + 
  geom_smooth(aes(colour = "lm"), method = "lm", se = FALSE) +
  labs(colour = "Method")
```